---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hottest100

<!-- badges: start -->
[![R-CMD-check](https://github.com/katerobsau/hottest100/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/katerobsau/hottest100/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/katerobsau/hottest100/graph/badge.svg)](https://app.codecov.io/gh/katerobsau/hottest100)
<!-- badges: end -->

The goal of hottest100 is to provide a compendium of Triple J hottest 100 songs. Data is scraped from [Triple J's best 100 Australian songs](https://www.abc.net.au/triplej/countdown/hottest100/australiansongs)
and historic information on [Wikipedia](https://en.wikipedia.org/wiki/Triple_J_Hottest_100).

There are three data sets on artists, tracks, and polls. 

## Installation

You can install the development version of hottest100 from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("katerobsau/hottest100")
```

## Data sets available

```{r}
library(hottest100)
library(tibble)
```

### Tracks

Triple J Hottest 100 Track Data - A dataset containing track-level data from the Triple J Hottest 100. A data.frame with one row per track, including variables:

* __alltime__: Logical; indicates if the entry is from an all-time chart specific year.
* __artist__: Character; name of the artist or band associated withtrack.
* __country__: Character; the country associated with the track/chart.
* __id__: Unique identifier for each track (within year)
* __pollyear__: Countdown poll year
* __position__: Chart position of the track.
* __track__: Title of the track or song.
* __releaseyear__: Original release year of the track.

```{r}
as_tibble(tracks)
```

### Polls

Triple J Hottest 100 Polls. A dataset with information about each Triple J Hottest 100 poll. 

A data.frame with one row per countdown, including variables:

* __year__: The year the poll was held.
* __alltime__: Logical; whether the poll was an "All-Time" countdown
* __alltimedescription__: A description of the poll, if applicable.
* __colorRGB__: Primary RGB color value associated with the year in character format.
* __contrastColorRGB__: Contrasting RGB color value (in character format) suitable for text overlay.

```{r}
as_tibble(polls)
```

### Artists

Triple J Hottest 100 Artists. A deduplicated dataset of artists represented in the Triple J Hottest 100, created from `tracks`.
 
* __artist__: Artist name
* __country__: Country of origin
* __releaseyear__: Release year of included track


```{r}
as_tibble(artists)
```
